---
output:
  word_document: default
  html_document: default
---
# Module 6- Assignment 2 
## Zeller, Ashton 
### Statistical Analysis 

```{r show_col_types = FALSE}
library(tidyverse)
library(readxl)
Copy_of_RespiratoryExchangeSample <- read_excel("Copy of RespiratoryExchangeSample.xlsx")
Copy_of_Perceptions <- read_excel("Copy of Perceptions.xlsx")
library(readr)
Advertising <- read_csv("Advertising.csv")
Insurance <- read_csv("Insurance.csv")

```

#### Regression and Correlation 
Regression analysis is a statistical method that allows you to examine the relationship between two or more 
variables of interest. Correlation analysis is a method of statistical evaluation used to study the strength of a 
relationship between two, numerically measured, continuous variables (e.g. height and weight). This 
particular type of analysis is useful when a researcher wants to establish if there are possible connections 
between variables. 

#### Insurance Costs
We would like to determine if we can accurately predict insurance costs based upon the factors included in 
the data. We would also like to know if there are any connections between variables (for example, is age 
connected or correlated to charges). 

#### Correlations of bmi, age, children and cost

```{r}
Insurance2 <- select(Insurance,"age","bmi","children","charges")
cor(Insurance2)
Corr_matrix <- cor(Insurance2)
library(corrplot) 
library(RColorBrewer) 
 
corrplot(Corr_matrix, type="upper", order="hclust", 
         col=brewer.pal(n=8, name="RdYlBu")) 
```

The graphic above is a correlation matrix. It shows what values are correlated. From this matrix above we can see that age and charges are highly correlated. 

#### Regression Analysis 
```{r}
Regression_Analysis <- lm(charges ~ age + bmi + children, data = Insurance2)
summary(Regression_Analysis)
```

Based on the results above, there is a high significance (.001) that our x-variable of age & bmi can predict our y-variable of charges.

```{r}
Insurance <- mutate(Insurance, gender=ifelse(sex=="female",1,0)) 
Insurance <- mutate(Insurance, smoker2=ifelse(smoker=="yes",1,0))
RA2 <- lm(charges ~ age + bmi + children + gender + smoker2, data = Insurance)
summary(RA2)
```

To see if smoking and gender had an impact on cost, I had to first convert those values to a zero or 1. With female = 1 and male = 0, non-smoker = 0, smoker = 1. Based on the results above, there is a high significance (.001) that our x-variable of smoker2 can predict our y-variable of charges. Meaning being a smoker has an impact on cost. 

#### Group Comparisons with t-tests 

 The t-test is used to compare the values of the means from two samples and test whether it is likely that the 
samples are from populations having different mean values. This is often used to compare 2 groups to see if 
there are any significant differences between these groups. 

#### Caffeine Impacts on Respiratory Exchange Ratio 

A study of the effect of caffeine on muscle metabolism used volunteers who each underwent arm exercise 
tests. Half the participants were randomly selected to take a capsule containing pure caffeine one hour 
before the test. The other participants received a placebo capsule. During each exercise the subject's 
respiratory exchange ratio (RER) was measured. (RER is the ratio of CO2 produced to O2 consumed and is an 
indicator of whether energy is being obtained from carbohydrates or fats). 
The question you are trying to answer is whether caffeine impacts RER during exercise. 

```{r}
summary(Copy_of_RespiratoryExchangeSample)

t.test(Copy_of_RespiratoryExchangeSample$Placebo,Copy_of_RespiratoryExchangeSample$Caffeine)
```

Based on the results of the t-test above, since the p-value is less than 0.05, it can be concluded that caffeine does not impact RER during exercise. 

#### Impact of Advertising 

You are a marketing researcher conducting a study to understand the impact of a new marketing campaign. 
To test the new advertisements, you conduct a study to understand how consumers will respond based on 
see the new ad compared to the previous campaign. One group will see the new ad and one group will see 
the older ads. They will then rate the ad on a scale of 0 to 100 as a percentage of purchase likelihood based 
on the ad. 
The question you are trying to answer is whether to roll out the new campaign or stick with the current 
campaign. 

```{r}
summary(Advertising)

t.test(Advertising$Group,Advertising$Rating,var.equal=TRUE)
```

Based on the results of the student t-test above, since the p-value is less than 0.05, it can be concluded to no tmove forward with the new marketing campaign. 

#### ANOVA
An ANOVA test is a way to find out if survey or experiment results are significant. In other words, they help 
you to figure out if you need to reject the null hypothesis or accept the alternate hypothesis. Basically, you’re 
testing groups to see if there’s a difference between them. Examples of when you might want to test 
different groups: 
A group of psychiatric patients are trying three different therapies: counseling, medication and 
biofeedback. You want to see if one therapy is better than the others. 
- A manufacturer has two different processes to make light bulbs. They want to know if one process is 
better than the other. 
- Students from different colleges take the same exam. You want to see if one college outperforms the 
other. 

#### Perceptions of Social Media Profiles 
This study examines how certain information presented on a social media site might influence perceptions of 
trust, connectedness and knowledge of the profile owner. Specifically, participants were shown weak, 
average and strong arguments that would influence their perceptions of the above variables. Using the 
dataset provided, the following code runs an ANOVA with post-hoc analyses to understand argument 
strength impacts on perceptions. 

```{r}
ANOVA1 <- aov(Trust ~ Argument, data=Copy_of_Perceptions)
summary(ANOVA1)
ANOVA2 <- aov(Connectedness ~ Argument, data = Copy_of_Perceptions)
summary(ANOVA2)
ANOVA3 <- aov(Knowledge ~ Argument, data = Copy_of_Perceptions)
summary(ANOVA3)

TukeyHSD(ANOVA1)
TukeyHSD(ANOVA2)
```

Based on the 3 ANOVA tests that were preformed above, two of the ANOVA's Pr(<F) were significant. Trust and Connectedness were significant across Argument. When conducting a Tukey Analysis to see the difference between a strong and average argument, we can see the ANOVA on Argument across Trust & Connectedness is not significant as p adj is 0.97(Trust) & 0.36 (Connectedness). To be significant it should be 0.05 or less. 




 
