---
output:
  word_document: default
  html_document: default
---
# Module 6 - Assignment 1 
## Zeller, Ashton
### Lubridate

```{r message="FASLE"}
library(tidyverse)
library(lubridate)
library(readr)
Appointments <- read_csv("Appointments.csv")
```

#### Working with Dates
```{r}
Appointments <- mutate(Appointments, ReservationYear =2019)
Appointments <- mutate(Appointments, ReservationDate =make_date(month=ReservationMonth,day=ReservationDay,year=ReservationYear))
Appointments <- mutate(Appointments,CreationDate =make_date(month=CreationMonth,day=CreationDay,year=CreationYear))
Appointments <- mutate(Appointments, ReservationSpan = ReservationDate-CreationDate)
```

#### Exploratory Analysis
```{r}
summary(Appointments)
Appointments$ReservationSpan <- as.numeric(Appointments$ReservationSpan)
summary(Appointments)
```
a). What is the average time between appointment creation and actual reservation date? 
Around 38 days. 
b). What was the max time? 
207 days. 
c). What was the min time? 
One day. 

```{r}
cor(Appointments$ReservationSpan,Appointments$Show)
```

a). Are ReservationSpan and Show highly correlated? 
ReservationSpan and Show are positively correlated, but not highly correlated. 

```{r}
Appointments[] <- lapply(Appointments, as.logical)
ggplot(data=Appointments, aes(x=Show))+
  geom_bar()
```

a). Did most people make their appointments? 
Yes, the majority of people made their appointments


